---
title: "US Immigration Analysis"
author: "Tsu-Hao Fu"
output:
  pdf_document: default
date: "2024-3-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(forecast)) install.packages("forecast")
library(forecast)
if(!require(tseries)) install.packages("tseries")
library(tseries)
```
## 1. Summary
This work identifies trends in the annual immigration levels in the United 
States. As part of the analysis, the study also identifies a model that best 
fits the immigrant data and forecasts the future number of immigrants entering 
the U.S. over the next eight years.

The data contains the total number of U.S. immigrants for each year ranging from 
1820 to 1920. Immigration has been increasing over time. In the early 1800s, it 
was low, but by the late 1800s, the U.S. had relaxed immigration policies. The 
number of immigrants peaked at the beginning of the 20th century, then decreased 
in the early 1900s due to World War I and laws in 1921 and 1924 that limited the
number of immigrants. Immigration dropped further during the Great Depression 
and World War II. After World War II, around 1950, it began to rise again.

```{r, warning=FALSE, echo = FALSE}
# Load the data
"us.immig" <-
  structure(c(8385, 9127, 6911, 6354, 7912, 10199, 10837, 18875, 
              27382, 22520, 23322, 22633, 48386, 58640, 65365, 45374, 76242, 
              79340, 38914, 68069, 84066, 80289, 104565, 69994, 78615, 114371, 
              154416, 234968, 226527, 297024, 295984, 379466, 371603, 368645, 
              427833, 200877, 200436, 251306, 123126, 121282, 153640, 91918, 
              91985, 176282, 193418, 248120, 318568, 315722, 277680, 352768, 
              387203, 321350, 404806, 459803, 313339, 227498, 169986, 141857, 
              138469, 177826, 457257, 669431, 788992, 603322, 518592, 395346, 
              334203, 490109, 546889, 444427, 455302, 560319, 579663, 439730, 
              285631, 258536, 343267, 230832, 229299, 311715, 448572, 487918, 
              648743, 857046, 812870, 1026499, 1100735, 1285349, 782870, 751786, 
              1041570, 878587, 838172, 1197892, 1218480, 326700, 298826, 295403, 
              110618, 141132, 430001, 805228, 309556, 522919, 706896, 294314, 
              304488, 335175, 307255, 279678, 241700, 97139, 35576, 23068, 
              29470, 34956, 36329, 50244, 67895, 82998, 70756, 51776, 28781, 
              23725, 28551, 38119, 108721, 147292, 170570, 188317, 249187, 
              205717, 265520, 170434, 208177, 237790, 321625, 326867, 253265, 
              260686, 265398, 271344, 283763), .Tsp = c(1820, 1962, 1), 
            class = "ts")

plot(us.immig, xlab="Year", ylab="Number of Immigrants", 
     main="Annual Immigration to the U.S. (1820-1962)",
      col="blue")
```

The immigration data seem to show there is a trend in the data. The 
trend seems to be a deterministic trend. In order to forecast future immigration 
levels, the trend needs to be removed from the series to achieve stationarity. 
By the Box-Cox transformation, log transformation should be used to stabilize a 
possible non-constant variance before taking a difference for stationary 
processes. An approximation of normality will also be improved by the log 
transformation. The plot of the transformed series shows an approximately 
increasing trend and strong momentum between observations which suggest 
non-stationarity. The observed increasing trend recommend taking a difference 
of the log-transformed data for stationarity. 

```{r, echo=FALSE}
# Box-Cox Transformation
cat("Lambda for Box-Cox:",BoxCox.lambda(us.immig))
us.immig = log(us.immig)
plot(us.immig, xlab="Year", ylab="Log Transformed Immigration Data", 
     main="Log Annual Immigration to the U.S. (1820-1962)",
     col="blue")
```

Our final ARIMA model:
$\Delta Z_t = -0.9831 \Delta Z_{t-1} - 0.6232 \Delta Z_{t-2} + 1.2334 a_{t-1} + 
0.7619 a_{t-2} + 0.1209 a_{t-3} + 0.0495 a_{t-4} - 0.1557 a_{t-5} - 
0.3493 a_{t-6} + a_t$

Our final ARI model:
$\Delta Z_t = 0.1833 \Delta Z_{t-1} - 0.1375 \Delta Z_{t-2} + 
0.2158 \Delta Z_{t-3} - 0.1399 \Delta Z_{t-4} - 0.1915 \Delta Z_{t-5} - 0.0794 
\Delta Z_{t-6} + 0.1658 \Delta Z_{t-7} + a_t
$

where \( \Delta Z_t \) is the first-differenced time series at time t and 
\( a_t \) is the white noise at time t.


## 2. Analysis
## 2.1 Stationarity Testing
By employing the Augmented Dickey-Fuller test, we observe that after the first 
difference is taken, the p-value falls below the significance level of 0.05. 
This suggests that the time series does not have a unit root, thereby indicating 
its stationarity. From the plot, the log-transformed immigration data also 
appears to be stationary after the first differencing.

```{r, warning=FALSE, echo=FALSE}
# Augmented Dickey-Fuller test
adf.test(us.immig)

# First difference
cat("Number of differences required to achieve stationarity:",ndiffs(us.immig),
    "\n")
diff_series <- diff(us.immig, differences=1)

# Augmented Dickey-Fuller test for the first difference
adf.test(diff_series)

# Plot the differenced series
plot(diff_series, main="First Differenced Series", 
     ylab="Differenced Immigration",
     col="blue")

```

## 2.2 Model Specification
The use of Autocorrelation Function (ACF) and Partial Autocorrelation Function 
(PACF) correlograms helped choose the ARIMA model parameters (p, q) for the data. 
The observations within the 95% significance bounds generally indicate a white 
noise process, with an exception of a notable spike at lag 5. This spike could be 
random or signify an important feature. Therefore, we assume that the optimal p 
and q are around 5.


```{r, echo=FALSE}
# ACF and PACF plots
acf(diff_series, main="ACF of Differenced Series")
pacf(diff_series, main="PACF of Differenced Series")
```

```{r, echo=FALSE}
# Grid Search function for the optimal ARIMA model
grid_search <- function(p_start=0,p_max=7,d = 1,q_start=0,q_max=7){
  
  best_aic <- Inf
  best_order <- c(0, d, 0)
  best_model <- NULL
  
  # Initialize a matrix to store the AIC values
  aic_matrix <- matrix(NA, nrow = p_max-p_start+1, ncol = q_max-q_start+1, 
                  dimnames = list(p = c(p_start:p_max), q = c(q_start:q_max)))
  i = 1
  for (p in p_start:p_max) {
    j = 1
    for (q in q_start:q_max) {
        # Fit the ARIMA model
        model <- arima(us.immig, order=c(p, d, q), method="ML")
        aic_matrix[i, j] <- model$aic
        
        # Check if this model has better AIC
        if (model$aic < best_aic) {
          best_aic <- model$aic
          best_order <- c(p, d, q)
          best_model <- model
        }
        j = j+1
    }
    i = i+1
  }
  # Print the AIC matrix
  print(aic_matrix)
  cat("The Order of the Optimal Model:",best_order)
  return(best_model)
}
```

## 2.3 ARIMA 
In order to select an appropriate ARIMA model that best fits the immigration 
data, a matrix of AIC values was used. The AIC values from matrix suggested 
ARIMA models such as ARIMA (2,1,6), ARIMA (3,1,6), and ARIMA (2,1,7). All three 
models were fit to the log transformed immigration data. The ARMA (2,1,6) model 
appeared to most adequately fit the immigration data.

The standardized residuals plot for the ARIMA (2,1,6) model indicates the 
residuals are homoscedastic and center around zero mean. The ACF and PACF 
correlograms show that the residuals appear to look like white noise. All of 
the p-values for the Ljung-Box statistic are above 0.05 at lags 1-20,suggesting 
the residuals do not show significant autocorrelation, and our model has 
adequately captured the autocorrelations in the data. Even though the normal 
Q-Q plot depict heavier tails indicating some skewness, overall, the ARIMA 
(2,1,6) model seem to have taken care of the significant spikes at lag 5.


```{r, warning=FALSE, echo=FALSE}
# ARIMA
best_ARIMA <- grid_search()
summary(best_ARIMA)
residuals_ARIMA <- scale(best_ARIMA$residuals)

# Check if residuals behave like white noise
plot(residuals_ARIMA, xlab="Year", 
     main="Standarized ARIMA Residuals",
     ylab="",
     col="blue")
```

```{r, warning=FALSE, echo=FALSE}
# Q-Q plot
qqnorm(residuals_ARIMA, main="Normal Q-Q Plot for Std Residuals (ARIMA)")
qqline(residuals_ARIMA)

# ACF and PACF
acf(best_ARIMA$residuals, main="ACF of ARIMA Residuals")
pacf(best_ARIMA$residuals, main="PACF of ARIMA Residuals")

# Ljung-Box test
p_values <- sapply(1:20, function(lag) {
  Box.test(best_ARIMA$residuals, lag = lag, type = "Ljung-Box")$p.value
})

# Plot the p-values for each lag
plot(1:20, p_values, type="p", lwd = 2, col = "blue", 
     xlab = "Lag", ylab = "P-Value", main = "Ljung-Box Test P-Values (ARIMA)")
abline(h = 0.05, col = "red", lty = 2)  # Reference line for p-value = 0.05


```

## 2.4 ARI
We follow the same procedures with ARIMA to find the best fitted model. The AIC 
values suggested ARI (7,1,0) is the most suitable model for the immigration data.

The standardized residuals plot for the ARI (7,1,0) model shows the residuals
are homoscedastic and centered around a zero mean, indicating a good fit. The 
ACF and PACF correlograms suggest the residuals resemble white noise, and all 
Ljung-Box test p-values are above 0.05, pointing to minimal autocorrelation. 
Although the normal Q-Q plot reveals some skewness due to heavier tails, it 
doesn't detract from the model's adequacy. Moreover, the ARI (7,1,0) model's "p"
is consistent with our initial assumption (p) based on the ACF correlogram for 
the first-differenced series. However, the ARIMA (2,1,6) is still a better fit 
to the data.

```{r, warning=FALSE, echo=FALSE}
# ARI
best_ARI <- grid_search(q_max=0)
summary(best_ARI)
residuals_ARI <- scale(best_ARI$residuals)
# Check if residuals behave like white noise
plot(residuals_ARI, xlab="Year", 
     main="Standarized Residuals (ARI)",
     col="blue")

# Q-Q plot
qqnorm(residuals_ARI, main="Normal Q-Q Plot for Std Residuals (ARI)")
qqline(residuals_ARI)

# ACF and PACF
acf(best_ARI$residuals, main="ACF of ARI Residuals")
pacf(best_ARI$residuals, main="PACF of ARI Residuals")

# Ljung-Box test
p_values <- sapply(1:20, function(lag) {
  Box.test(best_ARI$residuals, lag = lag, type = "Ljung-Box")$p.value
})

# Plot the p-values for each lag
plot(1:20, p_values, type="p", lwd = 2, col = "blue", 
     xlab = "Lag", ylab = "P-Value", main = "Ljung-Box Test P-Values (ARI)")
abline(h = 0.05, col = "red", lty = 2)  # Reference line for p-value = 0.05


```

## 2.5 Forcasting
The ARIMA (2,1,6) model was used to predict immigration levels for the years 
1963 to 1970. The plots also include the 80% and 95% confidence regions. 
Therefore, the predicted values show that immigration will gradually keep 
increasing for the next eight years.

```{r, warning=FALSE, echo=FALSE}
future_forecast <- forecast(best_ARIMA, h=8)
plot(future_forecast)
# Invert the log transformation for the forecast and its confidence intervals
original_scale_forecast <- exp(future_forecast$mean)
lower_ci <- exp(future_forecast$lower)
upper_ci <- exp(future_forecast$upper)
```

```{r, warning=FALSE, echo=FALSE}
# Plot the original non-transformed time series data
plot(exp(us.immig), type='l', main='Forecast on Original Scale', 
     ylab='Number of Immigrants', xlab='Year')

# Overlay the inverted forecast
lines(1963:1970, original_scale_forecast, col='blue', lwd=2)

# Add the confidence intervals
matlines(1963:1970, cbind(lower_ci[,1], upper_ci[,1]), col='red', lty=2, lwd=2)
matlines(1963:1970, cbind(lower_ci[,2], upper_ci[,2]), col='green', lty=2, lwd=2)

# Optionally, add a legend
legend("topleft", legend=c("Original Data", "Forecast", "80%CI", "95% CI"), 
       col=c("black", "blue", "red", "green"), lwd=2, lty=c(1, 1, 2, 2))

```